---
title: "Fitting models with purrr and broom"
author: "Casey Bates"
date: '2019-02-03'
slug: fitting-models-with-purrr-and-broom
tags:
- tidyverse
- broom
- purrr
- linear regression
- ggplot2
categories: R
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6, fig.height = 3.7)
```


```{r data_processing1, message=FALSE, include=FALSE}
# This chunk includes the data processing steps to wrangle and clean the data
library(tidyverse) # For reading-in, wrangling, visualizing the data, and for pipe operator
library(readxl) # Tidyverse package for reading in Excel files
library(magrittr) # For the '%<>%' pipe operator

# Read-in townhomes and condo sales data
townhomes <- read_xlsx("2017-comparable-sales-condos-townhomes.xlsx") 

# Read-in single family home sales data
single_family <- read_xlsx("2017-comparable-sales-single-family.xlsx") 

# Check column names and formats
glimpse(townhomes) 
glimpse(single_family)
# 'sale_price' column was coerced to numeric and special characters '$' and ',' removed

# Replace spaces in column names with underscores ('_') and make names lowercase
# the package 'magrittr' is used for the pipe operator, this is loaded with the 'tidyverse' library
colnames(townhomes) %<>% str_replace_all("\\s", "_") %<>% tolower() 
colnames(single_family) %<>% str_replace_all("\\s", "_") %<>% tolower()

# Rename column "actual_year_built" as "year_built" for consistency between datasets
townhomes <- townhomes %>% 
  rename(year_built = actual_year_built) 

# Add a "classifcation" column to the single_family dataset and set all values to "Single Family"
single_family <- single_family %>% 
  mutate(classification = "Single Family")

# bind_rows() throws an error b/c 'reception' column is numeric in townhomes dataset; convert to character
townhomes$reception <- as.character(townhomes$reception)

# Combine the datasets into one
home_sales <- bind_rows(single_family, townhomes)
glimpse(home_sales)
# No new columns were created during the 'bind_rows()' process, indicating all colums align
# Remove single_family and townhomes dataframes because they are no longer needed
rm(single_family, townhomes)

# unique(home_sales$classification) reveals a "Garage Only" type. Drop this.
home_sales <- home_sales %>% filter(classification != "Garage Only") 

# Rename the column "heated area" to "square_feet" for clarity, though technically not correct
home_sales <- home_sales %>% 
  rename(square_feet = heated_area)
```
```{r data_processing, message=FALSE, include=FALSE}
# This chunk includes the data processing steps to wrangle and clean the data
library(tidyverse) # For reading-in, wrangling, visualizing the data, and for pipe operator
library(readxl) # Tidyverse package for reading in Excel files
library(magrittr) # For the '%<>%' pipe operator

# Read-in townhomes and condo sales data
townhomes <- read_xlsx("2017-comparable-sales-condos-townhomes.xlsx") 

# Read-in single family home sales data
single_family <- read_xlsx("2017-comparable-sales-single-family.xlsx") 

# Check column names and formats
glimpse(townhomes) 
glimpse(single_family)
# 'sale_price' column was coerced to numeric and special characters '$' and ',' removed

# Replace spaces in column names with underscores ('_') and make names lowercase
# the package 'magrittr' is used for the pipe operator, this is loaded with the 'tidyverse' library
colnames(townhomes) %<>% str_replace_all("\\s", "_") %<>% tolower() 
colnames(single_family) %<>% str_replace_all("\\s", "_") %<>% tolower()

# Rename column "actual_year_built" as "year_built" for consistency between datasets
townhomes <- townhomes %>% 
  rename(year_built = actual_year_built) 

# Add a "classifcation" column to the single_family dataset and set all values to "Single Family"
single_family <- single_family %>% 
  mutate(classification = "Single Family")

# bind_rows() throws an error b/c 'reception' column is numeric in townhomes dataset; convert to character
townhomes$reception <- as.character(townhomes$reception)

# Combine the datasets into one
home_sales <- bind_rows(single_family, townhomes)
glimpse(home_sales)
# No new columns were created during the 'bind_rows()' process, indicating all colums align
# Remove single_family and townhomes dataframes because they are no longer needed
rm(single_family, townhomes)

# unique(home_sales$classification) reveals a "Garage Only" type. Drop this.
home_sales <- home_sales %>% filter(classification != "Garage Only") 

# Rename the column "heated area" to "square_feet" for clarity, though technically not correct
home_sales <- home_sales %>% 
  rename(square_feet = heated_area)
```


## Motivation

Utilize `tidyverse` tools and the `broom` package to generate numerous linear models to evaluate how well sale price is explained by square footage for home sales in Garfield County, Colorado. A dataset is publically available on the [Garfield County Assessor website](https://www.garfield-county.com/assessor/comparable-sales.aspx) that contains 2 years of data from summer 2014 through summer 2016.

## Processing the data
See the first blog post for all steps taken to clean the data. Here's a quick summary of the required data cleaning steps:

* Import two Excel files: 

  1. single family home sales, and 
  1. condo & townhome sales
* Replace spaces in column names with underscore and make lowercase
* Rename some columns
* Add `classification` column to `single_family` dataset
    * Set all values to "Single Family"
* Use `bind_rows()` to combine the datasets into one
* Remove "Garage Only" observations

## Glimpse of the data
```{r glimpse}
glimpse(home_sales)
```

## Many models with `purrr` and `broom`

## Remove erroneous high sale price observations
```{r remove_highs2}
home_sales_fix <- home_sales %>% filter(location != "RIFLE" | classification != "Townhome" | sale_price != 2200000)
home_sales_fix %>% filter(location == "RIFLE" & classification == "Townhome") %>% arrange(desc(sale_price)) %>% head(10)
```

## Linear model for Carbondale
```{r carbondale_lm}
carbondale <- home_sales_fix %>% 
  filter(location == "CARBONDALE")
model <- lm(sale_price ~ square_feet, data = carbondale)
summary(model)
```

## Transform model into data frame with `broom`
```{r tidy_model}
library(broom)
tidy(model)
```

## `bind_rows` to combine models
```{r two_models}
carbondale <- home_sales_fix %>% 
  filter(location == "CARBONDALE")
model1 <- lm(sale_price ~ square_feet, data = carbondale)
glenwood <- home_sales_fix %>% 
  filter(location == "GLENWOOD")
model2 <- lm(sale_price ~ square_feet, data = glenwood)
bind_rows(tidy(model1), tidy(model2))
```

## Step 1: `nest` by location
```{r step_1}
home_sales_fix_models <- home_sales_fix %>% 
  nest(-location)
home_sales_fix_models
```

## Alternate syntax used by Wickham
```{r alt_nest}
home_sales_fix %>% 
  group_by(location) %>% 
  nest()
```


## Examine the data for New Castle
```{r new_castle}
home_sales_fix_models$data[[3]]
```

## Unnesting returns to original
```{r unnest}
unnest(home_sales_fix_models)
```

## Step 2: `map()` to fit lm to each dataset
```{r step_2}
library(purrr)
home_sales_fix_models <- home_sales_fix %>%
  nest(-location) %>% 
  mutate(models = map(data, ~lm(sale_price ~ square_feet, .))) 
  # data has been passed into lm through map function
  # dot "." is used for data in the lm call
home_sales_fix_models
```

## Examine the model for New Castle
```{r new_castle_v2}
home_sales_fix_models$models[[3]]
```

## Step 3: Use `map()` to tidy each model
```{r step_3}
home_sales_fix_models <- home_sales_fix %>%
  nest(-location) %>% 
  mutate(models = map(data, ~lm(sale_price ~ square_feet, .))) %>% 
  mutate(tidied = map(models, tidy))
home_sales_fix_models
```

## Examine tidy model for New Castle
```{r new_castle_tidy}
home_sales_fix_models$tidied[[3]]
```

## Step 4: `unnest` to tidy table of coefficients
```{r step_4}
location_coeffs <- home_sales_fix %>%
  nest(-location) %>% 
  mutate(models = map(data, ~lm(sale_price ~ square_feet, .))) %>% 
  mutate(tidied = map(models, tidy)) %>% 
  unnest(tidied)
location_coeffs
```

## Location slopes
```{r slopes}
location_slopes <-  location_coeffs %>% 
  filter(term == "square_feet")%>% 
  arrange(estimate)
location_slopes
```

## Compute residuals for Rifle, pre-fix
```{r rifle_resids, message = FALSE}
library(modelr) # for add_residuals function
rifle <- filter(home_sales, location == "RIFLE")
rifle_model <- lm(sale_price ~ square_feet, data = rifle)
rifle %>% 
  add_residuals(rifle_model) %>% 
  ggplot(aes(square_feet, resid)) + # new resid column
  geom_hline(yintercept = 0, color = "white", size = 3) +
  geom_line() +
  ggtitle("Remaining pattern")
```

## Compute residuals for Rifle, fixed
```{r rifle_resids_v2}
rifle <- filter(home_sales_fix, location == "RIFLE")
rifle_model <- lm(sale_price ~ square_feet, data = rifle)
rifle %>% 
  add_residuals(rifle_model) %>% 
  ggplot(aes(square_feet, resid)) +
  geom_hline(yintercept = 0, color = "white", size = 3) +
  geom_line() +
  ggtitle("Remaining pattern")
```

## Compute residuals 
```{r add_resids, message=FALSE}
# Call add_residuals() with each model-data pair:
home_sales_fix_models <- home_sales_fix_models %>% 
  # Add 'resid' column to 'data' df list with mutate 
  mutate(data = map2(data, models, add_residuals))
home_sales_fix_models
```

## Examine residuals for New Castle
```{r new_castle_resids}
home_sales_fix_models$data[[3]] %>% 
  select(sale_price, square_feet, resid)
```

## `unnest` list of dataframes to plot resids
```{r unnest_resids, warning= FALSE}
resids <- unnest(home_sales_fix_models, data)
resids %>% 
  ggplot(aes(square_feet, resid)) +
  geom_line(aes(group = location, color=location), alpha = 1 / 3) +
  geom_smooth(se = FALSE)
```

## Or facet by location
```{r resids_facet_fix, warning= FALSE}
resids %>% 
  ggplot(aes(square_feet, resid, color = location)) +
  geom_line(alpha = 1 / 3) +
  facet_wrap(~ location)
```

## Pipe it all together!
```{r resids facet, warning= FALSE}
home_sales %>% # Original dataset for Rifle with errors!
  nest(-location) %>% 
  mutate(models = map(data, ~lm(sale_price ~ square_feet, .))) %>%
  mutate(data = map2(data, models, add_residuals)) %>%
  unnest(data) %>% 
  ggplot(aes(square_feet, resid, color = location)) +
  geom_line(alpha = 1 / 3) +
  facet_wrap(~ location)
```


